---
title: "**Run time analysis of Fibonacci Sequence**"
author: "Lakshman Madhav Kollipara"
date: "July 28, 2015"
output: html_document
---


The Fibonacci Sequence is the series of numbers: 0, 1, 1, 2, 3, 5, 8, 13, 21, 34, ...  
The next number is found by adding up the two numbers before it.  
The 2 is found by adding the two numbers before it (1+1).  
Similarly, the 3 is found by adding the two numbers before it (1+2),  
And the 5 is (2+3),  
and so on!  

I am doing Runtime analysis on Fibonacci Sequence using 4 different Algorithms.  
1. Algorithm 1: Dyanamic Programming[using Arrays], Complexity = O(n)  
2. Algorithm 2: Space optimized Dyanamic Programming [using temporary variables], Complexity = O(n)  
3. Algorithm 3: Recursive with Memoization[storing results of sub problems in global variable], Complexity = O(n)  
4. Algorithm 4: Recursive, Complexity = O(2^n)  


I calculated run time of each algorithm with various input sizes varying from 100 to 10000 using microbenchmark package.  
I have plotted the log-log plot on runtime data with Input size on X-axis and Runtime on Y-axis.  
Note:There are two separate plots; one for Fibonacci with recursive and other algorithms in another plot.   



**Description of R functions used:**  
**group_by()** - takes dataframe and grouping column as input and returns a data frame grouped by passed column.  
**do()** - Takes grouped data, Expression/function and optional paramenters as arguments, applies function to each group of passed data and stores results in a new column of Dataframe.   
**microbenchmark()** - Takes a set of functions, unit of time and number of times as input and runs those functions for specified number of times and returns runtimes of each run, which are organized as min,max, mean,lq,uq e.t.c.   
**ggplot2()** - Used to plot the graph and used geom_smooth to draw a regression curve.    

```{r}
rm(list = ls())##Clearing Environment

##Fibonacci_Iterative_using Array
fib_array <- function(n){
  fibvals <- numeric(n)
  if(n == 1){
    fibvals[1] <- 0
  }else if(n ==2){
    fibvals[1] <- 0
    fibvals[2] <- 1
  }else{
    fibvals[1] <- 0
    fibvals[2] <- 1
    for (i in 3:n) { 
      fibvals[i] <- fibvals[i-1]+fibvals[i-2]
    }
  }
  return(fibvals)
}

##Fibonacci_using Recursion
fib_recursive <- function(n){
  fibvals <- numeric(n)
  fib_recursive_cal <- function(n){ 
    if (n == 1) return(0) 
    if (n == 2) return(1) 
    return (fib_recursive_cal(n - 1) + fib_recursive_cal(n - 2)) 
  }
  for(i in 1:n)
  {
    fibvals[i] <- fib_recursive_cal(i)
  } 
  return(fibvals)
}

##Fibonacci using Recursion and Memoization
##Golbal vector to store results of sub problem
storage <- numeric(10000)
storage[1:length(storage)] <- NA
fib_recursive_mem <- function(n){
  temp <- numeric()
  if(is.na(storage[n])){ 
    if (n == 1) {
      storage[n] <<- 0
      return(0)
    } else if (n == 2){
      storage[1] <<- 0
      storage[n] <<- 1
      return(1)
    } 
    storage[n] <<- (fib_recursive_mem(n - 1) + fib_recursive_mem(n - 2)) 
    #print(storage[!is.na(storage)])
  }
  return(storage[n])
}

##fib_Iterative_without_array
fib_iterative <-function(n){
  first <- 0
  second <- 1
  fibvals <- numeric(n)
  for(i in 1:n){
    if(i == 1){
      next_e <- 0
    }
    else if(i ==2){
      next_e <- 1
    }
    else{
      next_e <- first + second
      first <- second
      second <- next_e
    }
    fibvals[i] <- next_e 
  }
  return(fibvals)
}


library(microbenchmark)
elapsed_time <- function(data_by_input, fib_array, fib_iterative, fib_recursive_mem){
  n <- data_by_input$input
  elapsed <- microbenchmark(fib_array(n),fib_iterative(n),fib_recursive_mem(n),unit = "ms", times = 1)
  ret <- data.frame(run_time = elapsed$time, algorithm = as.character(elapsed$expr))
  return(ret)
}
library(dplyr)
run_time <- data.frame(input = seq(from = 100, to = 10000, by = 100))
data_by_input <- group_by(run_time, input)
runtime_data <- do(data_by_input,elapsed_time(., fib_array,fib_iterative,fib_recursive_mem))
#print(runtime_data)
runtime_log_data <- data.frame(input= log10(runtime_data$input), run_time = log10(runtime_data$run_time), algorithm = runtime_data$algorithm)
#print(runtime_log_data)


##Separate plot for fib_recursive because of different scale from rest of the plots

elapsed_time_rec <- function(data_by_input_rec, fib_recursive){
  n <- data_by_input_rec$input
  elapsed_rec <- microbenchmark(fib_recursive(n),unit = "ms", times = 1)
  ret_rec <- data.frame(run_time = elapsed_rec$time, algorithm = as.character(elapsed_rec$expr))
  return(ret_rec)
}
run_time_rec <- data.frame(input = seq(from = 5, to = 30, by = 5))
data_by_input_rec <- group_by(run_time_rec, input)
runtime_data_rec <- do(data_by_input_rec,elapsed_time_rec(., fib_recursive))

```


```{r, echo=FALSE}
library(ggplot2)
p <- ggplot(runtime_log_data) + geom_point(aes(x = input, y = run_time, color = algorithm)) + 
  geom_smooth(aes(x = input, y = run_time, color = algorithm), method= "lm") 
plot(p)
p_rec <- ggplot(runtime_data_rec) + geom_point(aes(x = input, y = run_time, color = algorithm)) + 
  geom_smooth(aes(x = input, y = run_time, color = algorithm), method= "loess", span = 1) 
plot(p_rec)
```
